---
title: "Practical Machine Learning - Prediction Assignment Writeup"
author: "New Yi Qing, Isaac"
date: "Sunday, March 22, 2015"
output: html_document
---
```{r, echo=FALSE, message=FALSE, warning = FALSE, error = FALSE}
library(knitr)
```
  
###Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

###Data
The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment. 

###What you should submit
The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases. 

1. Your submission should consist of a link to a Github repo with your R markdown and compiled HTML file describing your analysis. Please constrain the text of the writeup to < 2000 words and the number of figures to be less than 5. It will make it easier for the graders if you submit a repo with a gh-pages branch so the HTML page can be viewed online (and you always want to make it easy on graders :-).
2. You should also apply your machine learning algorithm to the 20 test cases available in the test data above. Please submit your predictions in appropriate format to the programming assignment for automated grading. See the programming assignment for additional details. 

###Reproducibility
Due to security concerns with the exchange of R code, your code will not be run during the evaluation by your classmates. Please be sure that if they download the repo, they will be able to view the compiled HTML version of your analysis.

###Loading Required Libraries
```{r, echo=TRUE}
library(caret)
library(rpart)
library(randomForest)
```

###Loading Dataset
```{r, echo=TRUE}
naStrings <- c("NA","#DIV/0!","")
training <- read.csv("data/pml-training.csv", na.strings=naStrings)
testing <- read.csv("data/pml-testing.csv", na.strings=naStrings)
```

###Data Processing
Dataset is cleaned from variables with near zero variance
```{r, echo=TRUE}
nzvTraining <- nearZeroVar(training, saveMetrics = TRUE)
sum(nzvTraining$nzv)
```
36 variables with near zero variance are removed
```{r, echo=TRUE}
training <- training[, !nzvTraining$nzv]
```
Dataset is cleaned from Variables with more than 60% missing values
```{r, echo=TRUE}
na60Percent <- sapply(colnames(training), function(x) 
    if(sum(is.na(training[, x])) > 0.60*nrow(training)) {
        return(TRUE)
    }else{
        return(FALSE)
    }
)
sum(na60Percent)
```
65 variables with more than 60% missing values are removed
```{r, echo=TRUE}
training <- training[, !na60Percent]
```
Dataset is cleaned from variables related with data acquisition (e.g id, timestamps, individuals' names, etc.) which are not suitable to be used in prediction
```{r, echo=TRUE}
training <- training[, -(1:6)]
```
Correlation analysis
```{r, echo=TRUE}
corr <- caret::findCorrelation(cor(training[, -53]), cutoff=0.8)
names(training)[corr]
```
Many variables are highly correlated. PCA will be used in the pre-processing. After the data cleaning, the variables that are selected to specify the model are:
```{r, echo=TRUE}
names(training)
```
###Partioning the training set into two for training and testing
60% for myTraining, 40% for myTesting
```{r, echo=TRUE}
inTrain <- createDataPartition(y=training$classe, p=0.6, list=FALSE)
myTraining <- training[inTrain, ] 
myTesting <- training[-inTrain, ]
dim(training)
dim(myTraining) 
dim(myTesting)
```

###Using Decision Tree algorithms for prediction 
```{r, echo=TRUE}
modelRPART <- rpart(classe ~ ., data=myTraining, method="class")
```
Predicting
```{r, echo=TRUE}
predictionsRPART <- predict(modelRPART, myTesting, type = "class")
```
Using confusion matrix to test results
```{r, echo=TRUE}
confusionMatrix(predictionsRPART, myTesting$classe)
```
###Using Random Forests algorithms for prediction
```{r, echo=TRUE}
modelRF <- randomForest(classe ~. , data=myTraining)
```
Predicting
```{r, echo=TRUE}
predictionsRF <- predict(modelRF, myTesting, type = "class")
```
Using confusion matrix to test results
```{r, echo=TRUE}
confusionMatrix(predictionsRF, myTesting$classe)
```
###Summary
Random Forests yielded better results in terms of accuracy and Kappa.
  
###Generating files for assignment submission
Using Random Forests prediction for assignment submission
```{r, echo=TRUE}
predictionsRF <- predict(modelRF, testing, type = "class")
pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

pml_write_files(predictionsRF)
```
  
~~~ End ~~~